---
title: "ggiraph R package"
output:
  github_document:
    toc: true 
    toc_depth: 2
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

> Make 'ggplot' Graphics Interactive

[![Travis-CI Build Status](https://travis-ci.org/davidgohel/ggiraph.svg?branch=master)](https://travis-ci.org/davidgohel/ggiraph)
[![version](http://www.r-pkg.org/badges/version/ggiraph)](http://www.r-pkg.org/pkg/ggiraph)
![cranlogs](http://cranlogs.r-pkg.org./badges/ggiraph) 
![Active](http://www.repostatus.org/badges/latest/active.svg)

## Overview

`ggiraph` is an [htmlwidget](http://cran.r-project.org/package=htmlwidgets) and a ggplot2 extension. It lets ggplot graphics being animated. 

Animation is made with ggplot geometries that can understand three arguments:

* `tooltip`: column of dataset that contains tooltips to be displayed when mouse is over elements. 
* `onclick`: column of dataset that contains javascript function to be executed when elements are clicked. 
* `data_id`: column of dataset that contains id to be associated with elements. 

If used within a shiny application, elements associated with an id (`data_id`) can be selected and manipulated on client and server sides.

```{r results='asis', echo=FALSE}
htmltools::HTML(sprintf('<img src="data:image/gif;base64,%s"/>', 
        gdtools:::base64_file_encode(filename = "/Users/davidgohel/Github/ggiraph/inst/ggiraph.gif")
        ))

```

## Installation 

##### github

    # get development version
    devtools::install_github('davidgohel/ggiraph')

##### cran

    # get cran version
    install.packages("ggiraph")


