---
title: "ggiraph R package"
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

[![Build Status](https://travis-ci.org/davidgohel/ggiraph.svg?branch=master)](https://travis-ci.org/davidgohel/ggiraph)
[![Build status](https://ci.appveyor.com/api/projects/status/github/davidgohel/ggiraph?branch=master)](https://ci.appveyor.com/project/davidgohel/ggiraph/branch/master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/ggiraph)](https://cran.r-project.org/package=ggiraph)
![cranlogs](http://cranlogs.r-pkg.org./badges/ggiraph)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)

<p align="center">
  <img width="15%" src="https://www.ardata.fr/img/hexbin/ggiraph.svg">
</p>

> Make 'ggplot' Graphics Interactive

## Overview

`ggiraph` is an htmlwidget and a ggplot2 extension. It allows ggplot
graphics to be interactive, by exporting them as SVG documents and using 
special attributes on the various elements.

Interactivity is added to ggplot geometries, legends and theme elements, via 
some parameters:

* `tooltip`: column of dataset that contains tooltips to be displayed when mouse is over elements. 
* `onclick`: column of dataset that contains javascript function to be executed when elements are clicked. 
* `data_id`: column of dataset that contains id to be associated with elements. 

If used within a shiny application, elements associated with an id (`data_id`) 
can be selected and manipulated on client and server sides.

<iframe width="100%" height="550" src="https://www.youtube.com/embed/cJt5hlCi_do" frameborder="0" allowfullscreen></iframe>



## Installation 

##### Get development version on github

```{r eval=FALSE}
devtools::install_github('davidgohel/ggiraph')
```

##### Get CRAN version

```{r eval=FALSE}
install.packages("ggiraph")
```    

