---
title: "ggiraph R package"
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

> Make 'ggplot' Graphics Interactive

[![Travis-CI Build Status](https://travis-ci.org/davidgohel/ggiraph.svg?branch=master)](https://travis-ci.org/davidgohel/ggiraph)
[![Build status](https://ci.appveyor.com/api/projects/status/github/davidgohel/ggiraph?branch=master)](https://ci.appveyor.com/project/davidgohel/ggiraph/branch/master)
[![version](http://www.r-pkg.org/badges/version/ggiraph)](https://CRAN.R-project.org/package=ggiraph)
![cranlogs](http://cranlogs.r-pkg.org./badges/ggiraph)
![Active](http://www.repostatus.org/badges/latest/active.svg)

## Overview

`ggiraph` is an htmlwidget and a ggplot2 extension. It lets ggplot graphics being animated. 

Animation is made with ggplot geometries that can understand three arguments:

* `tooltip`: column of dataset that contains tooltips to be displayed when mouse is over elements. 
* `onclick`: column of dataset that contains javascript function to be executed when elements are clicked. 
* `data_id`: column of dataset that contains id to be associated with elements. 

If used within a shiny application, elements associated with an id (`data_id`) can be selected and manipulated on client and server sides.

```{r results='asis', echo=FALSE}
htmltools::HTML(sprintf('<img src="data:image/gif;base64,%s"/>',
        gdtools:::base64_file_encode(filename = "/Users/davidgohel/Github/ggiraph/inst/ggiraph.gif")
        ))
```

## Installation 

##### Get development version on github

```{r eval=FALSE}
devtools::install_github('davidgohel/ggiraph')
```

##### Get CRAN version

```{r eval=FALSE}
install.packages("ggiraph")
```    


