---
title: "ggiraph R package"
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<p align="center">
  <img width="15%" src="https://www.ardata.fr/img/hexbin/ggiraph.svg">
</p>

> Make 'ggplot' Graphics Interactive

## Overview

`ggiraph` is an htmlwidget and a ggplot2 extension. It allows ggplot graphics to be animated. 

Animation is made with ggplot geometries that can understand three arguments:

* `tooltip`: column of dataset that contains tooltips to be displayed when mouse is over elements. 
* `onclick`: column of dataset that contains javascript function to be executed when elements are clicked. 
* `data_id`: column of dataset that contains id to be associated with elements. 

If used within a shiny application, elements associated with an id (`data_id`) 
can be selected and manipulated on client and server sides.

<iframe width="100%" height="550" src="https://www.youtube.com/embed/cJt5hlCi_do" frameborder="0" allowfullscreen></iframe>



## Installation 

##### Get development version on github

```{r eval=FALSE}
devtools::install_github('davidgohel/ggiraph')
```

##### Get CRAN version

```{r eval=FALSE}
install.packages("ggiraph")
```    

