# Advanced customization

{{< include _init.qmd >}}

```{r}
p1 <- ggplot(
  data = diamonds, 
  mapping = aes(x = color, fill = cut, data_id = cut)) +
  geom_bar_interactive(
    aes(tooltip = sprintf("%s: %.0f", fill, after_stat(count))),
    size = 3)
```

## Controling svg identifier

It is possible to control the value of the SVG ID produced. In this case, the
value must be assigned to the `canvas_id` argument in the call to `girafe()`
which will be passed to the `ggiraph::dsvg()` function.

```{r}
girafe(ggobj = p1, canvas_id = "myid")
```

## Adding attributes to interactive elements

Adding attributes to interactive elements is an advanced option that can be used to add a class name or any attribute.

```{r}
p1 <- ggplot(
  data = diamonds, 
  mapping = aes(x = color, fill = cut, data_id = cut)) +
  geom_bar_interactive(
    aes(tooltip = sprintf("%s: %.0f", fill, after_stat(count))),
    size = 3, `data-blah`="myclass", extra_interactive_params = c("data-blah"),
    onclick = "alert(this.getAttribute(\"data-blah\"));")
girafe(ggobj = p1)
```

