---
title: "A quick introduction"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{A quick introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The **ggiraph** package let R users make **ggplot interactive**. The package is an htmlwidget. Below is an example:

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(ggiraph)
theme_set(theme_minimal())

# geom_point_interactive example
gg_point_1 <- ggplot(mtcars, aes(x = disp, y = qsec, 
		color = wt, tooltip = row.names(mtcars), data_id = row.names(mtcars) ) ) + 
	geom_point_interactive(size=3) + 
  scale_color_gradient(low = "#F3C899", high = "#8C120A") 

# htmlwidget call
ggiraph(code = {print(gg_point_1)}, width_svg = 5, height_svg = 4, width = .75,
        tooltip_extra_css = "padding:2px;background:rgba(70,70,70,0.1);color:black;border-radius:2px 2px 2px 2px;",
        hover_css = "fill:#1279BF;stroke:#1279BF;cursor:pointer;", pointsize = 14)

```

It extends ggplot2 with new `geom` functions:

```{r echo=FALSE, results='asis'}
geoms_ <- objects(envir = as.environment("package:ggiraph"), pattern = "^geom_([a-zA-Z0-9]*)_interactive$")
htmltools::tags$ul( lapply( geoms_, htmltools::tags$li ) )
```

These understand three aesthetics to let you add interactivity:

* `tooltip`: column of dataset that contains tooltips to be displayed when mouse is over elements. 
* `onclick`: column of dataset that contains javascript function to be executed when elements are clicked. 
* `data_id`: column of dataset that contains id to be associated with elements. **This aesthetic is mandatory 
  when you want to use an hover effect or when you want to enable selection of points in shiny applications.**

Let's prepare a ggplot object with the `mpg` dataset. 

```{r}
library(ggplot2)
library(ggiraph)

head(mpg)
g <- ggplot(mpg, aes( x = displ, y = cty, color = hwy) )
```

## tooltips

The first example shows how to add a tooltip:

```{r message=FALSE}
my_gg <- g + geom_point_interactive(aes(tooltip = model), size = 2) 
ggiraph(code = print(my_gg) )
```

## hover effects

Now let's add an hover effect. **Elements associated with a `data_id` will be animated upon mouse over.**

```{r message=FALSE}
my_gg <- g + geom_point_interactive(
    aes(tooltip = model, data_id = model), size = 2) 
ggiraph(code = print(my_gg), hover_css = "cursor:pointer;fill:red;stroke:red;")
```

The default value of hover css is `hover_css = "fill:orange;"`. To see how to change that, read the *custome_effects* vignette.

> Note that `data-id` can also be reused within a shiny application. 

## Click actions

### Using onclick

Click actions must be a string column in the dataset containing valid javascript instructions. 

```{r message=FALSE, warning=FALSE}
crimes <- data.frame(state = tolower(rownames(USArrests)), USArrests)
head(crimes)

# create an 'onclick' column
crimes$onclick <- sprintf("window.open(\"%s%s\")",
  "http://en.wikipedia.org/wiki/", as.character(crimes$state) )

gg_crime <- ggplot(crimes, aes(x = Murder, y = Assault, color = UrbanPop )) + 
  geom_point_interactive(
    aes( data_id = state, tooltip = state, onclick = onclick ), size = 3 ) + 
  scale_colour_gradient(low = "#999999", high = "#FF3333")

ggiraph(code = print(gg_crime), hover_css = "fill-opacity:.3;cursor:pointer;")
```

### within shiny

When working with shiny, you can use the `data_id` aesthetic to associate points, polygons and other graphical elements with a value that will be available in a reactive context. This makes it possible to click on an element and trigger an action. More details are available in vignette "ggiraph in shiny". Note that in this case, `onclick` should not be used, both onclick and data_id will need the "click" event.
