---
title: "Custom animation effects"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Custom animation effects}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

With `ggiraph`, you can customize tooltip style and mouse hover effects. This requires usage of `css`.

## Tooltip position

The arguments `tooltip_offx` and `tooltip_offy` are used to offset tooltip position. 

By default the offset is 10 pixels horizontally to the mouse position (`tooltip_offx=10`) and 0 pixels vertically (`tooltip_offx=10`).

```{r message=FALSE}
library(ggiraph)
theme_set(theme_minimal())

dataset <- mtcars
dataset$carname <- row.names(dataset)
gg_point_1 <- ggplot(dataset, aes(x = disp, y = qsec, tooltip = carname, data_id = carname, color= wt) ) + 
	geom_point_interactive(size=3)

# htmlwidget call
ggiraph(code = {print(gg_point_1)}, tooltip_offx = 20, tooltip_offy = -10 )
```


## Tooltip style

The `ggiraph` function has an argument named `tooltip_extra_css`. It can be used to add css declarations to customize tooltip rendering.

> Each css declaration includes a property name and an associated value. Property names and values are separated by colons and name-value pairs always end with a semicolon. For example `color:gray;text-align:center;`. Common properties are : 

> * background-color: background color 
* color: elements color
* border-style, border-width, border-color: border properties
* width, height: size of tooltip
* padding: space around content

Tooltip opacity can be defined with the argument `tooltip_opacity` (default to 0.9).

Let's custom tooltip as:

* italic font 
* no background color

```{r}
tooltip_css <- "background-color:transparent;font-style:italic;"
```

Now print the ggiraph:

```{r message=FALSE}
ggiraph(code = {print(gg_point_1)}, tooltip_extra_css = tooltip_css )
```

Now, let's add a gray rectangle with round borders and a few other details to make it less crude:

```{r}
tooltip_css <- "background-color:gray;color:white;font-style:italic;padding:10px;border-radius:10px 20px 10px 20px;"

ggiraph(code = {print(gg_point_1)}, tooltip_extra_css = tooltip_css, tooltip_opacity = .75 )
```

> Do not surround `tooltip_extra_css` value by curly braces, ggiraph takes care of that.

## Hover effects

Hover effects occur when the mouse is **over elements that have a `data-id` attribute** (resulting from using argument `data_id` in interactive geom functions). It will only modify SVG element rendering when the mouse is over an element. 

Mouse over effects can be configured with the `hover_css` argument in the same way `tooltip_extra_css` is used for customizing tooltip rendering.

css here is relative to SVG elements. SVG attributes are listed [here](https://developer.mozilla.org/fr/docs/Web/SVG/Attribute). Common properties are: 

* fill: background color 
* stroke: color
* stroke-width: border width
* r: circle radius (**no effect if Firefox is used**).

To fill elements in red:

```{r}
ggiraph(code = {print(gg_point_1)}, hover_css = "fill:red;r:10pt;" )
```

## Zoom

You can activate zoom; set zoom_max (maximum zoom factor) to a value greater than 1.

```{r}
ggiraph(code = print(gg_point_1), zoom_max = 5)
```


