---
title: "ggiraph presentation"
author: "David Gohel"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The **ggiraph** package let R users to make **ggplot interactive**. The package is an [htmlwidget](http://cran.r-project.org/package=htmlwidgets). Below an example:

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(ggiraph)

# geom_point_interactive example
gg_point_1 <- ggplot(mtcars, aes(x = disp, y = qsec, 
		color = wt, tooltip = row.names(mtcars), data_id = row.names(mtcars) ) ) + 
	geom_point_interactive(size=3) + 
  scale_color_gradient(low = "#F3C899", high = "#8C120A") 

gg_point_1 <- gg_point_1 + theme_minimal()
# htmlwidget call
ggiraph(code = {print(gg_point_1)}, width = .6,
        tooltip_extra_css = "padding:2px;background:rgba(70,70,70,0.1);color:black;border-radius:2px 2px 2px 2px;",
        hover_css = "fill:#1279BF;stroke:#1279BF;cursor:pointer;")

```

It extends ggplot2 with new `geom` functions:

* geom_bar_interactive
* geom_point_interactive
* geom_polygon_interactive
* geom_map_interactive
* geom_path_interactive
* geom_rect_interactive
* geom_segment_interactive
* geom_text_interactive


Three arguments let you add interactivity:

* `tooltip`: column of dataset that contains tooltips to be displayed when mouse is over elements. 
* `onclick`: column of dataset that contains javascript function to be executed when elements are clicked. 
* `data_id`: column of dataset that contains id to be associated with elements. 

Let's prepare a ggplot object with `mpg` dataset. 

```{r}
library(ggiraph)

head(mpg)
g <- ggplot(mpg, aes( x = displ, y = cty, color = drv) ) + theme_minimal()
```

## tooltips

The first example show how to add tooltip:

```{r message=FALSE}
my_gg <- g + geom_point_interactive(aes(tooltip = model), size = 2) 
ggiraph(code = print(my_gg), width = .7)
```

## hover effects

Now let's add an hover effect. **Elements associated with a `data_id` will be animated when mouse will be hover.**

```{r message=FALSE}
my_gg <- g + geom_point_interactive(
    aes(tooltip = model, data_id = model), size = 2) 
ggiraph(code = print(my_gg), width = .7)
```

Default value of hover css is `hover_css = "fill:orange;"`. To see how to change that, read the *custome_effects* vignette.

> Note that `data-id` can also be reused within a shiny application. 

## Click actions

Click actions must be a string column in the dataset containing valid javascript instructions. 

```{r message=FALSE, warning=FALSE}
crimes <- data.frame(state = tolower(rownames(USArrests)), USArrests)
head(crimes)

# create an 'onclick' column
crimes$onclick <- sprintf("window.open(\"%s%s\")",
  "http://en.wikipedia.org/wiki/", as.character(crimes$state) )

gg_crime <- ggplot(crimes, aes(x = Murder, y = Assault, color = UrbanPop )) + 
  geom_point_interactive(
    aes( data_id = state, tooltip = state, onclick = onclick ), size = 3 ) + 
  scale_colour_gradient(low = "#999999", high = "#FF3333") + 
  theme_minimal()

ggiraph(code = print(gg_crime),
        hover_css = "fill-opacity:.3;cursor:pointer;")
```


## Zoom 

You can activate zoom; set `zoom_max` (maximum zoom factor) to a value greater than 1. 

```{r}
ggiraph(code = print(my_gg + theme_linedraw()), zoom_max = 5)
```

