---
title: "ggiraph"
author: "David Gohel"
date: "2015/08/29"
output: html_document
---

The **ggiraph** package let R users to make **ggplot interactive**. The format used is SVG. 
The package is an [htmlwidget](https://cran.r-project.org/web/packages/htmlwidgets/index.html). 
For now, there are three arguments to let you add interactivity:

* `tooltip`: column of dataset that contains tooltips to be displayed when mouse is over elements. 
* `onclick`: column of dataset that contains javascript function to be executed when elements are clicked. 
* `data_id`: column of dataset that contains id to be associated with elements. 

```{r echo=FALSE, warning=FALSE,error=FALSE, message=FALSE, prompt=FALSE}
library(ggiraph)
```

# Simple example 
```{r}
dataset = iris
dataset$tooltip = paste0( "Species: ", dataset$Species )
dataset$clickjs = paste0("function() {alert('",dataset$Species, "')}" )

# plots
gg_point_1 = ggplot(dataset, aes(x = Sepal.Length, y = Petal.Width, 
		color = Species, tooltip = tooltip, onclick=clickjs) ) + 
	geom_point_interactive(size=3)

ggiraph(fun=print, x = gg_point_1)
```

# Installation


You will need the latest `ggplot2` version (>= 2.0.0) and package `rvg`.

    install.packages(c("ggplot2", "rvg"))
    devtools::install_github('davidgohel/ggiraph')
