---
title: "ggiraph and Shiny"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ggiraph and Shiny}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`ggiraph` is an htmlwidget. It can be used within a Shiny application. The package contains two Shiny examples. You can run them and see the code by running:

```{r eval=FALSE}
shiny::runApp(appDir = system.file("shiny/crimes", package = "ggiraph"), display.mode = "showcase")
shiny::runApp(appDir = system.file("shiny/cars", package = "ggiraph"), display.mode = "showcase")
```


## The client `ui.R`

Instead of a `plotOutput`, use function `ggiraphOutput`. 

```{r eval=FALSE}
ggiraphOutput("plot")
```

## The server `server.R`

Use function `ggiraphOutput`.

```{r eval=FALSE}
output$plot <- renderggiraph({
    ggiraph(code = print(gg_blahblah) )
  })
```

## Selections

It is possible to work with selecteds point on a *ggiraph* plot within a Shiny application.

Selection can be of two types: `single` or `multiple`. The ggiraph's parameter `selection_type` will let you specify that. 

```{r eval=FALSE}
output$myplot <- renderggiraph({
    ggiraph(code = print(gg_blahblah), selection_type = "multiple" )
  })
```

The selected points will be captured in the input reactive value `myplot_selected` (name of the input id of the reactive output value + `_selected`):

```{r eval=FALSE}
input$myplot_selected
```


You can also modify theses values by using the `session$sendCustomMessage` method with type `myplot_set` (name of the input id of the reactive output value + `_set`).

```{r eval=FALSE}
# delete selection
session$sendCustomMessage(type = 'myplot_set', message = character(0))
```

