---
title: "ggiraph in shiny"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{ggiraph in shiny}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

`ggiraph` is an htmlwidget. It can be used within a Shiny application. In shiny, elements associated with `data_id` can be selected and the selection (the `data_id` value) is available in the client and the server side of the application. 

## The client `ui.R`

Instead of a `plotOutput`, use function `ggiraphOutput`. 

```{r eval=FALSE}
ggiraphOutput("plot")
```

## The server `server.R`

Use function `ggiraphOutput`.

```{r eval=FALSE}
output$plot <- renderggiraph({
    ggiraph(code = print(gg_blahblah) )
  })
```

## Selections

It is possible to work with selecteds point on a *ggiraph* plot within a Shiny application.

Selection can be of two types: `single` or `multiple`. The ggiraph's parameter `selection_type` will let you specify that. 

```{r eval=FALSE}
output$myplot <- renderggiraph({
    ggiraph(code = print(gg_blahblah), selection_type = "multiple" )
  })
```

The selected points will be captured in the input reactive value `myplot_selected` (name of the input id of the reactive output value + `_selected`):

```{r eval=FALSE}
input$myplot_selected
```


You can also modify theses values by using the `session$sendCustomMessage` method with type `myplot_set` (name of the input id of the reactive output value + `_set`).

```{r eval=FALSE}
# delete selection
session$sendCustomMessage(type = 'myplot_set', message = character(0))
```


## Examples

The package contains Shiny examples available in the `shiny` directory of the package (`system.file("shiny", package = "ggiraph")`).

`crimes` and `cars` applications demonstrate multiple and single selections usage.

```{r eval=FALSE}
shiny::runApp(appDir = system.file("shiny/crimes", package = "ggiraph"), display.mode = "showcase")
shiny::runApp(appDir = system.file("shiny/cars", package = "ggiraph"), display.mode = "showcase")
```

`DT` demonstrates how to use `onclick` actions on the client side.

```{r eval=FALSE}
shiny::runApp(appDir = system.file("shiny/DT", package = "ggiraph"), display.mode = "showcase")
```

`iris` demonstrates behavior of selections when some `data_id` are duplicated.

```{r eval=FALSE}
shiny::runApp(appDir = system.file("shiny/iris", package = "ggiraph"), display.mode = "showcase")
```

