---
title: "Customizing girafe objects"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Customizing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
library(knitr)
opts_chunk$set(message=FALSE)
```

You can customize tooltip style, mouse hover effects, toolbar position... 
This requires usage of `css` instructions and options. The function `girafe` has 
an argument `options`, a list of options to customize the rendering. 
These effects will be defined by function calls, i.e. `opts_tooltip()`, `opts_toolbar()`, ...

## Tooltip options

Tooltip can be customized with function `opts_tooltip()`.

### Tooltip position

The arguments `offx` and `offy` are used to offset tooltip position. 

By default the offset is 10 pixels horizontally to the mouse position (`offx=10`) and 0 
pixels vertically (`offy=0`).

```{r}
library(ggplot2)
library(ggiraph)

dataset <- mtcars
dataset$carname <- row.names(dataset)
gg_scatter <- ggplot(dataset, aes(x = disp, y = qsec, tooltip = carname, data_id = carname, color= wt) ) + 
	geom_point_interactive(size=3)

girafe(ggobj = gg_scatter, 
  options = list(opts_tooltip(offx = 20, offy = 20)) )
```

If argument `use_cursor_pos` is set to FALSE, the mouse position will not 
be used and the tooltip will be fixed at `offx` and `offy`.

```{r}
girafe(ggobj = gg_scatter, 
  options = list(
    opts_tooltip(offx = 60, offy = 60, use_cursor_pos = FALSE)) )
```

### Tooltip style

The `opts_tooltip` function has an argument named `css`. It can be used to add css declarations to customize tooltip rendering.

> Each css declaration includes a property name and an associated value. Property names and values are separated by colons and name-value pairs always end with a semicolon. For example `color:gray;text-align:center;`. Common properties are : 

> * background-color: background color 
* color: elements color
* border-style, border-width, border-color: border properties
* width, height: size of tooltip
* padding: space around content

Tooltip opacity can be defined with the argument `opacity` (default to 0.9).

Let's custom tooltip as:

* italic font 
* no background color

```{r}
tooltip_css <- "background-color:transparent;font-style:italic;"
```

Now print the girafe graphic:

```{r message=FALSE}
girafe(ggobj = gg_scatter, 
  options = list(opts_tooltip(css = tooltip_css) ) )
```

Now, let's add a gray rectangle with round borders and a few other details to make it less crude:

```{r}
tooltip_css <- "background-color:gray;color:white;font-style:italic;padding:10px;border-radius:5px;"
girafe(ggobj = gg_scatter, 
  options = list(opts_tooltip(css = tooltip_css) ) )
```

> Do not surround `css` value by curly braces, `girafe` function takes care of that.

### Auto coloring

In function `opts_tooltip`, set argument `use_fill` to TRUE and the background 
color of tooltip will always use use elements fill property to color tooltip.

Argument `use_stroke` is to be used to apply the same to the border color of the tooltip.

```{r}
girafe(ggobj = gg_scatter, 
  options = list(opts_tooltip(use_fill = TRUE) ) )
```


Package `ggiraph` enable elements to be dynamic when mouse is hovering over them. 
This is possible when an element is associated with a `data_id`.

The dynamic aspect of elements can be defined with css code by the user. There are several ways 
to define these settings. 

## Hover effects

The elements that are associated with a `data_id` are animated when the mouse 
hovers over them. These animations can be configured using the following functions:

* `opts_hover` for the animation of panel elements
* `opts_hover_key` for the animation of the elements of the legends
* `opts_hover_theme` for the animation of the elements of the theme

These functions all have a `css` argument that defines by css instructions 
the style to use when the mouse passes over them. `css` here 
is relative to SVG elements. SVG attributes are listed [here](https://developer.mozilla.org/fr/docs/Web/SVG/Attribute). Common 
properties are: 

* fill: background color 
* stroke: color
* stroke-width: border width
* r: circle radius (**no effect if Firefox is used**).

To fill elements in red, `opts_hover` call should be used as below: 

```{r}
girafe(ggobj = gg_scatter, 
  options = list(
    opts_hover(css = "fill:red;")
  ) )
```


### Detailled control

Now there are cases where css expressions will have to be configured with 
more caution. Let's have a look at the following example ; if you put your 
mouse hover points or text, you will see that the animation is not 
adapted to the text. Text should instead be animated with another css property. 


```{r}
gg_scatter2 <- ggplot(dataset, aes(x = disp, y = qsec, label = carname, 
  data_id = carname, color= wt) ) + 
  geom_point_interactive(size=3) + geom_text_interactive(vjust = 2)
girafe(ggobj = gg_scatter2, 
  options = list(
    opts_hover(css = "fill:red;stroke:black;")
  ) )
```

Function `girafe_css` is to be used in that case, it allows to specify 
individual styles for various SVG elements.

```{r}
girafe(ggobj = gg_scatter2, 
  options = list(
    opts_hover(
    css = girafe_css(
      css = "fill:purple;stroke:black;",
      text = "stroke:none;fill:red;"
    )
  )
  ) )
```


## Zoom

You can activate zoom; set `zoom_max` (maximum zoom factor) to a value 
greater than 1. If the argument is greater than 1, a toolbar will 
appear when mouse will be over the graphic. 

Click on the icons in the toolbar to activate or desactivate the zoom.

```{r}
girafe(ggobj = gg_scatter, 
  options = list(
    opts_zoom(max = 5))
  )
```

## Toolbar

A toolbar is added by default to the graphs at the top right. It contains 
at least the download button for the PNG version of the graph. 

It will contain other elements depending on the options used. If a zoom 
has been configured, the zoom options will be added to it. If a selection 
is configured and the graph is in 
a shiny application, the selection options will be added to it, i.e. selection 
and anti-selection. 

Toolbar position cann be defined with function `opts_toolbar()` and argument `position`.  

```{r}
girafe(ggobj = gg_scatter, 
  options = list(
    opts_toolbar(position = "bottomright") )
  )
```

Also 'save as png' button can be desactivated by using argument `saveaspng`.  
saveaspng relies on JavaScript promises, so any browsers that don't natively support the standard Promise object will need to have a polyfill (e.g. Internet Explorer with version less than 11 will need it).


```{r}
girafe(ggobj = gg_scatter, 
  options = list(
    opts_toolbar(saveaspng = FALSE) )
  )
```

## Size options

By default the size of the graph is automatically adjusted to 100% of the 
width of the web page containing it. 

> Graphic will be resized if its container is resized.

This behavior can be controlled by using the `opts_sizing` function.

The percentage of the width to be used can be defined with the `width` 
parameter which takes a value between 0 and 1. 

```{r}
girafe(ggobj = gg_scatter, 
  options = list(
    opts_sizing(rescale = TRUE, width = .7) )
  )
```

This resizing can also be undone using the `rescale` argument with the 
value `FALSE`. In this case, the use of `width` will have no effect.  
To control the actual size of the graph, use the `width_svg` and `height_svg` 
parameters of the `girafe` function, the expected values are expressed in inches.


```{r}
girafe(ggobj = gg_scatter, width_svg = 7, height_svg = 4, 
  options = list(
    opts_sizing(rescale = FALSE) )
  )
```



## Fonts 

When you use specific fonts, you will need that font installed on your system.
This can be check with package `gdtools` and function `gdtools::font_family_exists()`.

```{r}
gdtools::font_family_exists("Roboto")
```

To use that font in a *girafe* graphic, best is to use the argument `fonts` (see `dsvg()`).
It's a named list that can contains 4 font family names, one per family type: `sans`, 
`serif`, `mono` and `symbol`. 

The following example illustrates how to use font *Roboto* as *sans* family.

```{r eval=gdtools::font_family_exists("Roboto")}
z <- ggplot(data = iris, 
  mapping = aes(x = Sepal.Length, y = Petal.Width)) +
  geom_point() +
  theme_minimal(base_family = "sans", base_size = 18)

girafe(ggobj = z, fonts = list(sans = "Roboto"))
```

To be able to show the graphic on a web page, you must make sure 
the font will be available. To do so, one solution is to use 
css in the R Markdown document or in the shiny application. 

The following css code is generated by https://fonts.google.com/ 
and allows to embed the font `Roboto` in a web page.

```{css}
@font-face {
  font-family: 'Roboto';
  font-style: normal;
  font-weight: 400;
  font-display: swap;
  src: local('Roboto'), local('Roboto-Regular'), url(https://fonts.gstatic.com/s/roboto/v20/KFOmCnqEu92Fr1Mu72xKKTU1Kvnz.woff2) format('woff2');
  unicode-range: U+0460-052F, U+1C80-1C88, U+20B4, U+2DE0-2DFF, U+A640-A69F, U+FE2E-FE2F;
}
```

