---
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Arrange multiple plots into a grid 

There are two avalaible options to arrange ggiraph outputs into a grid. As 
`ggiraph` is mainly only new *geoms*, package `cowplot` and `patchwork` 
can be used seamlessly.

```{r message=FALSE}
library(ggplot2)
library(ggiraph)
library(patchwork)
library(cowplot)

mtcars <- mtcars
mtcars$tooltip <- rownames(mtcars)

theme_set(theme_minimal())

gg1 <- ggplot(mtcars) +
  geom_point_interactive(aes(x = drat, y = wt, color = qsec, 
    tooltip = tooltip, data_id = tooltip ), size = 4) 

gg2 <- ggplot(mtcars) +
  geom_point_interactive(aes(x = qsec, y = disp, color = mpg, 
    tooltip = tooltip, data_id = tooltip ), size = 4)
```


### Using `cowplot`

Mouseover points to see where they are located in the other graph.

```{r}
ggiraph( code = print(gg1 + gg2), width_svg = 8, height_svg = 4, 
         hover_css = "fill:red;r:10pt;")
```

### Using `patchwork`

Mouseover points to see where they are located in the other graph.

```{r}
ggiraph( code = print(gg1 + gg2), width_svg = 8, height_svg = 4, 
         hover_css = "fill:red;r:10pt;")
```

## Dynamic dendrogram

[expr.R](./expr.R)

```{r message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggdendro)
library(purrr)
library(ggiraph)

source("expr.R")
expr_set_extract[1:4, 4:9]
```

First compute data for dendrograms

```{r message=FALSE}
hc <- hclust(dist(expr_set_extract), "ave")
dhr <- as.dendrogram(hc)
order_r <- rownames(expr_set_extract)[hc$order]


hc <- hclust(dist(t(expr_set_extract)), "ave")
dhc <- as.dendrogram(hc)
order_c <- colnames(expr_set_extract)[hc$order]

data_r <- dendro_data(dhr, type = "rectangle")
data_r <- segment(data_r)
data_r <- data_r %>%
  mutate( x_ = y + length(order_c),
          xend_ = yend + length(order_c),
          y_ = x,
          yend_ = xend )
data_c <- dendro_data(dhc, type = "rectangle")
data_c <- segment(data_c) %>% mutate(
  y = y + length(order_r),
  yend = yend + length(order_r)
)
```

Create a data.frame from the matrix, then order. 

```{r message=FALSE}
expr_set <- bind_cols(
  data_frame(gene = rownames(expr_set_extract)),
  as.data.frame(expr_set_extract)
)
expr_set <- gather(expr_set, variable, measure, -gene)
expr_set$gene <- factor( expr_set$gene, levels = order_r )
expr_set$variable <- factor( expr_set$variable, levels = order_c )
expr_set <- arrange(expr_set, gene, variable)
```

Add a variable to have tooltips as tables

```{r}
str_model <- paste0("<tr><td>gene</td><td>%s</td></tr>", 
  "<tr><td>variable</td><td>%s</td></tr>", 
  "<tr><td>measure</td><td>%.03f</td></tr>")
expr_set$tooltip <- sprintf(str_model, expr_set$gene, 
                            expr_set$variable, expr_set$measure )
expr_set$tooltip <- paste0( "<table>", 
                            expr_set$tooltip, "</table>" )
head(expr_set)
```

Create the ggplot using `geom_tile_interactive`

```{r}
p <- ggplot(data = expr_set, aes(x = variable, y = gene) ) +
  geom_tile_interactive(aes(fill = measure, tooltip = tooltip), colour = "white") +
  scale_fill_gradient(low = "white", high = "#BC120A", limits = c(0, 13)) +
  geom_segment(
    data = data_c,
    mapping = aes(x = x, y = yend, xend = xend, yend = y),
    colour = "gray20", size = .2) +
  geom_segment(
    data = data_r,
    mapping = aes(x = x_, y = y_, xend = xend_, yend = yend_),
    colour = "gray20", size = .2) +
  coord_equal()
```

Theme the object

```{r}
p <- p + theme_minimal() +
  theme(
    legend.position = "right",
    panel.grid.minor = element_line(color = "transparent"),
    panel.grid.major = element_line(color = "transparent"),
    axis.ticks.length	= unit(2, units = "mm"),
    plot.title = element_text(face = "bold", hjust = 0.5, size = 12),
    axis.title = element_text(size = 9, colour = "gray30"),
    axis.text.y = element_text(hjust = 1, size = 5, colour = "gray40"),
    axis.text.x = element_text(angle = 90, hjust = 1, size = 5, colour = "gray40"),
    legend.title=element_text(face = "bold", hjust = 0.5, size=8),
    legend.text=element_text(size=6)
  )
```

Use `ggiraph`

```{r}
ggiraph(ggobj = p, width_svg = 8, height_svg = 6, zoom_max = 2)
```

## Network with `ggraph`

The following code is adapted from http://www.pieceofk.fr/?p=431. 

```{r message=FALSE}
library(tidyverse)
library(stringr)
library(igraph)
library(tidygraph)
library(ggraph)
library(magrittr)
library(ggiraph)

pdb <- tools::CRAN_package_db()

aut <- pdb$Author %>%
  str_replace_all("\\(([^)]+)\\)", "") %>%
  str_replace_all("\\[([^]]+)\\]", "") %>%
  str_replace_all("<([^>]+)>", "") %>%
  str_replace_all("\n", " ") %>%
  str_replace_all("[Cc]ontribution.* from|[Cc]ontribution.* by|[Cc]ontributors", " ") %>%
  str_replace_all("\\(|\\)|\\[|\\]", " ") %>%
  iconv(to = "ASCII//TRANSLIT") %>%
  str_replace_all("'$|^'", "") %>%
  gsub("([A-Z])([A-Z]{1,})", "\\1\\L\\2", ., perl = TRUE) %>%
  gsub("\\b([A-Z]{1}) \\b", "\\1\\. ", .) 

aut <- aut %>%
  purrr::map(str_split, ",|;|&| \\. |--|(?<=[a-z])\\.| [Aa]nd | [Ww]ith | [Bb]y ", simplify = TRUE) %>%
  purrr::map(str_replace_all, "[[:space:]]+", " ") %>%
  purrr::map(str_replace_all, " $|^ | \\.", "") %>%
  purrr::map(function(x) x[str_length(x) != 0]) %>%
  set_names(pdb$Package) %>%
  extract(map_lgl(., function(x) length(x) > 1))

aut_list <- aut %>%
  unlist() %>%
  dplyr::as_data_frame() %>%
  count(value) %>%
  rename(Name = value, Package = n) %>% 
  mutate(Name = str_replace_all(Name, "'", " ")) 


edge_list <- aut %>%
  purrr::map(combn, m = 2) %>%
  do.call("cbind", .) %>%
  t() %>%
  dplyr::as_data_frame() %>%
  arrange(V1, V2) %>%
  count(V1, V2)

g <- edge_list %>%
  select(V1, V2) %>%
  as.matrix() %>%
  graph.edgelist(directed = FALSE) %>%
  as_tbl_graph() %>%
  activate("edges") %>%
  mutate(Weight = edge_list$n) %>%
  activate("nodes") %>%
  rename(Name = name) %>%
  mutate(Component = group_components()) %>%
  filter(Component == names(table(Component))[which.max(table(Component))])


g <- g %>%
  left_join(aut_list) %>%
  filter(Package > 4) %>%
  mutate(Component = group_components()) %>%
  filter(Component == names(table(Component))[which.max(table(Component))])


g <- mutate(g, Community = group_edge_betweenness(),
            Degree = centrality_degree())


g <- g %>%
  mutate(Community = case_when(Community == names(sort(table(Community),
                                                       decr = TRUE))[1] ~ "The Ancients",
                               Community == names(sort(table(Community),
                                                       decr = TRUE))[2] ~ "The Moderns",
                               Community %in% names(sort(table(Community),
                                                         decr = TRUE))[-1:-2] ~ "Unclassified")) %>%
  mutate(Community = factor(Community))

g <- g %>%
  filter(Degree > 5) %>%
  mutate(Degree = centrality_degree())
```

We can now create the plot :

```{r message=FALSE}
ggg <- ggraph(g, layout = 'lgl') +
  geom_edge_fan(alpha = 0.1, edge_width = .2) +
  geom_point_interactive(aes(x, y,
                             tooltip = Name, data_id = Name, color = Community,
                             size = Package), alpha = .7 ) +
  theme_graph() +
  scale_color_manual(values=c("The Ancients" = "#5BC8AC", "The Moderns" = "#F18D9E", Unclassified = "#E6D72A")) +
  theme(legend.position = "bottom")

tooltip_css <- "background-color:white;color:#333333;font-style:italic;padding:5px;border-radius:3px 4px;border: 1pt #333333 solid;"
hover_css <- "stroke:#333333;stroke-width:1pt;fill:red;cursor:pointer;"

ggiraph(ggobj = ggg, width_svg = 8, height_svg = 6,
        tooltip_extra_css = tooltip_css,
        hover_css = hover_css, width = 1, zoom_max = 4)
```




