---
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Arrange multiple plots into a grid 

There are two avalaible options to arrange ggiraph outputs into a grid. As 
`ggiraph` is mainly only new *geoms*, package `cowplot` and `patchwork` 
can be used seamlessly.

```{r message=FALSE}
library(ggplot2)
library(ggiraph)
library(patchwork)
library(cowplot)

mtcars <- mtcars
mtcars$tooltip <- rownames(mtcars)

theme_set(theme_minimal())

gg1 <- ggplot(mtcars) +
  geom_point_interactive(aes(x = drat, y = wt, color = qsec, 
    tooltip = tooltip, data_id = tooltip ), size = 4) 

gg2 <- ggplot(mtcars) +
  geom_point_interactive(aes(x = qsec, y = disp, color = mpg, 
    tooltip = tooltip, data_id = tooltip ), size = 4)
```


### Using `cowplot`

Mouseover points to see where they are located in the other graph.

```{r}
ggiraph( code = print(gg1 + gg2), width_svg = 8, height_svg = 4, 
         hover_css = "fill:red;r:10pt;")
```

### Using `patchwork`

Mouseover points to see where they are located in the other graph.

```{r}
ggiraph( code = print(gg1 + gg2), width_svg = 8, height_svg = 4, 
         hover_css = "fill:red;r:10pt;")
```

## Dynamic dendrogram

[expr.R](./expr.R)

```{r message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggdendro)
library(purrr)
library(ggiraph)

source("expr.R")
expr_set_extract[1:4, 4:9]
```

First compute data for dendrograms

```{r message=FALSE}
hc <- hclust(dist(expr_set_extract), "ave")
dhr <- as.dendrogram(hc)
order_r <- rownames(expr_set_extract)[hc$order]


hc <- hclust(dist(t(expr_set_extract)), "ave")
dhc <- as.dendrogram(hc)
order_c <- colnames(expr_set_extract)[hc$order]

data_r <- dendro_data(dhr, type = "rectangle")
data_r <- segment(data_r)
data_r <- data_r %>%
  mutate( x_ = y + length(order_c),
          xend_ = yend + length(order_c),
          y_ = x,
          yend_ = xend )
data_c <- dendro_data(dhc, type = "rectangle")
data_c <- segment(data_c) %>% mutate(
  y = y + length(order_r),
  yend = yend + length(order_r)
)
```

Create a data.frame from the matrix, then order. 

```{r message=FALSE}
expr_set <- bind_cols(
  data_frame(gene = rownames(expr_set_extract)),
  as.data.frame(expr_set_extract)
)
expr_set <- gather(expr_set, variable, measure, -gene)
expr_set$gene <- factor( expr_set$gene, levels = order_r )
expr_set$variable <- factor( expr_set$variable, levels = order_c )
expr_set <- arrange(expr_set, gene, variable)
```

Add a variable to have tooltips as tables

```{r}
str_model <- paste0("<tr><td>gene</td><td>%s</td></tr>", 
  "<tr><td>variable</td><td>%s</td></tr>", 
  "<tr><td>measure</td><td>%.03f</td></tr>")
expr_set$tooltip <- sprintf(str_model, expr_set$gene, 
                            expr_set$variable, expr_set$measure )
expr_set$tooltip <- paste0( "<table>", 
                            expr_set$tooltip, "</table>" )
head(expr_set)
```

Create the ggplot using `geom_tile_interactive`

```{r}
p <- ggplot(data = expr_set, aes(x = variable, y = gene) ) +
  geom_tile_interactive(aes(fill = measure, tooltip = tooltip), colour = "white") +
  scale_fill_gradient(low = "white", high = "#BC120A", limits = c(0, 13)) +
  geom_segment(
    data = data_c,
    mapping = aes(x = x, y = yend, xend = xend, yend = y),
    colour = "gray20", size = .2) +
  geom_segment(
    data = data_r,
    mapping = aes(x = x_, y = y_, xend = xend_, yend = yend_),
    colour = "gray20", size = .2) +
  coord_equal()
```

Theme the object

```{r}
p <- p + theme_minimal() +
  theme(
    legend.position = "right",
    panel.grid.minor = element_line(color = "transparent"),
    panel.grid.major = element_line(color = "transparent"),
    axis.ticks.length	= unit(2, units = "mm"),
    plot.title = element_text(face = "bold", hjust = 0.5, size = 12),
    axis.title = element_text(size = 9, colour = "gray30"),
    axis.text.y = element_text(hjust = 1, size = 5, colour = "gray40"),
    axis.text.x = element_text(angle = 90, hjust = 1, size = 5, colour = "gray40"),
    legend.title=element_text(face = "bold", hjust = 0.5, size=8),
    legend.text=element_text(size=6)
  )
```

Use `ggiraph`

```{r}
ggiraph(ggobj = p, width_svg = 7, height_svg = 7, zoom_max = 4)
```

