---
title: "examples"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Arrange multiple plots into a grid 

There are two avalaible options to arrange ggiraph outputs into a grid. As 
`ggiraph` is mainly only new *geoms*, package `cowplot` and `patchwork` 
can be used seamlessly.

```{r message=FALSE}
library(ggplot2)
library(ggiraph)
library(patchwork)
library(cowplot)

mtcars <- mtcars
mtcars$tooltip <- rownames(mtcars)

theme_set(theme_minimal())

gg1 <- ggplot(mtcars) +
  geom_point_interactive(aes(x = drat, y = wt, color = qsec, 
    tooltip = tooltip, data_id = tooltip ), size = 4) 

gg2 <- ggplot(mtcars) +
  geom_point_interactive(aes(x = qsec, y = disp, color = mpg, 
    tooltip = tooltip, data_id = tooltip ), size = 4)
```


### Using `cowplot`

Mouseover points to see where they are located in the other graph.

```{r}
girafe( ggobj = plot_grid(gg1, gg2), width_svg = 8, height_svg = 4)
```

### Using `patchwork`

Mouseover points to see where they are located in the other graph.

```{r}
girafe( code = print(gg1 + gg2), width_svg = 8, height_svg = 4)
```

## Inversed hovering

```{r}
library(ggplot2)
library(ggiraph)
library(data.table)
library(charlatan)

species <- ch_taxonomic_species(n = 10)
dat <- lapply(species, function(species, n){
  data.table(
    date = as.Date(seq_len(n), origin = "2018-10-25"),
    sales = cumsum(runif(n, -1, 1)),
    species = species,
    name = ch_name(n = n)
  )
}, n = 200)
dat <- rbindlist(dat)

gg <- ggplot(dat, aes(x = date, y = sales, 
                      colour = species, group = species)) +
  geom_line_interactive(aes(tooltip = name, data_id = species)) +
  scale_color_viridis_d() + 
  theme_minimal() + 
  labs(title = "move mouse over lines")

x <- girafe(ggobj = gg, width_svg = 8, height_svg = 6,
  options = list(
    opts_hover_inv(css = "opacity:0.1;"),
    opts_hover(css = "stroke-width:2;")
  ))
x
```



## With package `{ggraph}`

```{r message=FALSE}
library(ggraph)
library(igraph)
library(ggiraph)

actors <- data.frame(
  name = c("Alice", "Bob", "Cecil", "David", "Esmeralda"),
  age = c(48,33,45,34,21),
  gender = c("F","M","F","M","F")
)
relations <- data.frame(
  from = c("Bob", "Cecil", "Cecil", "David", "David", "Esmeralda"),
  to = c("Alice", "Bob", "Alice", "Alice", "Bob", "Alice"),
  same.dept = c(FALSE,FALSE,TRUE,FALSE,FALSE,TRUE),
  friendship = c(4,5,5,2,1,1), 
  advice = c(4,5,5,4,2,3)
)

g <- graph_from_data_frame(relations, 
       directed = TRUE, vertices = actors)

z <- ggraph(g, layout = 'linear', circular = TRUE) + 
  geom_edge_arc(color = "red", edge_width = .2) + 
  geom_point_interactive(size = 5,
    mapping = aes(x = x, y = y, data_id = gender,
                  tooltip = paste0(name, ": ", age, "y.o."))
  ) + theme_graph()

girafe(ggobj = z, width_svg = 5, height_svg = 5, 
       options = list(opts_sizing(rescale = FALSE)))

```

