---
title: "Fonts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fonts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
library(knitr)
library(htmltools)
library(ggplot2)
library(ggiraph)
opts_chunk$set(message=FALSE, eval=require("gdtools", quietly = TRUE))
```

```{r echo=FALSE}
if (require("gdtools", quietly = TRUE)) {
  register_gfont("Open Sans")

set_girafe_defaults(fonts = list(sans = "Open Sans"))

theme_set(
  theme_minimal(base_family = "Open Sans") +
    theme(
      plot.background	= element_blank(),
      panel.background = element_blank()
    )
)
}
addGFontHtmlDependency(family = "Open Sans")
```



```{r}
library(ggplot2)
library(ggiraph)
```

The ggiraph package generates SVG files and for this purpose, fonts are used, in particular to determine the size and position of graphic elements. For textual representations, the name of the font used is associated with the text. 

Be careful, if the font is not available on the machine where the image is displayed (the browser of one of your readers), the text will appear badly positioned.

You must therefore make sure that the font is available both on the computer that generates the SVG and on the computer used to view the SVG.


## Font managmement

If you do not specify the fonts to use, default fonts will be chosen for you. They are chosen based on the ones available on the machine. 

You can use the `validated_fonts()` function to see these values:

```{r}
validated_fonts()
```

In order to use a specific font, it must be available on your system. This can be checked with the `font_family_exists()` function.

```{r}
font_family_exists("Open Sans")
```

To take advantage of "Open Sans", it must be installed on the machine. 

You can also use fonts without installing them. You need to have the ttf files. For example, the package [gdtools](https://cran.r-project.org/package=gdtools) allows to use any 'Google Fonts', here the "Open Sans" font. 

```{r}
library(gdtools)
```

The following code allows to register a font that can then be used 
by 'ggiraph' graphic engine.

```{r eval=FALSE}
register_gfont("Open Sans")
```

The following code embeds the font in the HTML document and ensures that the
rendering will be possible for the reader even if he does not have the font
installed on his machine. If you want to do that by hand, you will need to learn
how to create a CSS "font-face".

```{r}
addGFontHtmlDependency(family = "Open Sans")
```

`register_gfont()` use `systemfonts::register_font()` to allow the use of a
font by ggiraph when creating the SVG file. `systemfonts` is can be used for font
management. Read systemfonts documentation if you need more details.

```{r eval=FALSE}
library(systemfonts)
if(!font_family_exists("Open Sans")){
  register_font(name = "Open Sans", 
    plain = list("fonts/open-sans-v18-latin-regular.woff", 0), 
    bold = list("fonts/open-sans-v18-latin-700.woff", 0), 
    italic = list("fonts/open-sans-v18-latin-italic.woff", 0),
    bolditalic = list("fonts/open-sans-v18-latin-700italic.woff", 0)
  )
}
```

## Font usage

To use that font in a *girafe* graphic, various options are available.

* You can use the argument `fonts` (see `?dsvg`). It's a named list that can contains 4 font family names, one per family type: `sans`, `serif`, `mono` and `symbol` where you should map with the font family names you'd like to use. 

```{r}
dat <- mtcars
dat$carname <- row.names(dat)

gg <- ggplot(dat, aes(drat, carname)) + 
  geom_point() + 
  theme_minimal(base_family = "sans")

validated_fonts(list(sans = "Open Sans"))

ggiraph(ggobj = gg, fonts = list(sans = "Open Sans"))
```

* Or you can simply use ggplot2 usual features:

```{r}

gg <- ggplot(dat, aes(drat, carname)) + 
  geom_point() + 
  theme_minimal(base_family = "Open Sans")

ggiraph(ggobj = gg)
```

## Complex example

Register the fonts with `add_font()`:

```{r}
register_gfont("Roboto")
register_gfont("Fira Code")
```

Add the fonts in the R markdown HTML result.

```{r}
addGFontHtmlDependency(family = c("Roboto", "Fira Code"))
```

Create the graphic and use fonts:

```{r}
gg <- ggplot(dat, aes(drat, carname)) + 
  geom_point() + 
  geom_text(data = dat[1:3,], aes(label = carname), color="orange", size = 5, family = "Fira Code")+
  theme_minimal(base_family = "sans") + 
  theme(axis.title = element_text(family = "Roboto"),
        axis.text.x = element_text(face = "italic"),
        axis.text.y = element_text(face = "bold")
        ) 

ggiraph(ggobj = gg, fonts = list(sans = "Open Sans"))
```

## Shiny illustration

Below an example of a simple shiny application that 
use a specific font and embed it in the application.

```{r eval=FALSE}
library(ggiraph)
library(ggplot2)
library(shiny)
library(gdtools)

register_gfont("Ms Madi")

ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      addGFontHtmlDependency(family = c("Ms Madi")),
      sliderInput("num_rows", "Number of rows:",
        min = 15, max = 32, value = 20)
    ),
    mainPanel(girafeOutput("custofontplot"))
  )
)

server <- function(input, output) {
  output$custofontplot <- renderGirafe({
    dat <- mtcars[seq_len(input$num_rows), ]
    dat$carname <- row.names(dat)
    gg <- ggplot(dat, aes(drat, carname)) +
      geom_point() +
      theme_minimal(base_family = "Ms Madi")

    girafe(ggobj = gg)
  })
}

# Run the application
shinyApp(ui = ui, server = server)
```

