---
title: "Shiny usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Shiny usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
library(knitr)
opts_chunk$set(message=FALSE)
```

A major advantage of `{ggiraph}` used in a shiny application is the ability to
collect selections.

The user defines them by clicking on the elements or by defining an area
surrounding the elements to be selected (lasso selection). It is also possible
to deselect all or part of an existing selection.

## Shiny usage 

You will need to use `girafeOutput` in the UI part of the application and 
`renderGirafe` in the SERVER part.

> The client `ui.R`

Instead of a `plotOutput`, use function `girafeOutput`. 

```{r eval=FALSE}
girafeOutput("plot")
```

> The server `server.R`

Use function `renderGirafe`.

```{r eval=FALSE}
output$plot <- renderGirafe({
    girafe(ggobj = gg_blahblah )
  })
```

## Examples

The package contains Shiny examples available in the `shiny` 
directory of the package (`system.file("shiny", package = "ggiraph")`).

Use function `run_girafe_example` to launch the applications 
in showcase mode. They can be used to better understand how 
to use girafe graphics with shiny applications. 

The following applications are available: `cars`, `click_scale`, 
`crimes`, `DT`, `dynamic_ui`, `iris`, `gender`, `maps` and `modal`.

### Panel selections usage (server side)

```{r eval=FALSE}
run_girafe_example("crimes")
```

### Legend selections usage (server side)

```{r eval=FALSE}
run_girafe_example("click_scale")
```

### onclick actions (client side).

```{r eval=FALSE}
run_girafe_example("DT")
```

### Girafe reactivity showoff

```{r eval=FALSE}
run_girafe_example("gender")
```

## Working with user selections

The graphics produced by `girafe` from a shiny application will allow you to
retrieve the element selections made by users.

Elements associated with `data_id` can be selected and the selection (the
`data_id` value) is available in the client and the server side of the
application. **The selected identifiers will be the values mapped by the
aesthetic `data_id`**.

The selection type can take several values: `single`, `multiple` or `none`. 

- `single`: the user can only select one element. The click allows its
selection if it is not selected, or its de-selection if it is already
selected. Clicking on an unselected element automatically de-selects the
other selected element.
- `multiple`: the user can select several elements. He can do this by clicking
on the elements or by selecting in the toolbar the "lasso selection" menu
which allows you to draw a lasso on the graph and select all the elements
contained in the lasso. The toolbar also contains an "anti-lasso selection"
menu that allows you to draw a lasso on the graph and de-select all the
elements contained in the lasso. The click is of course available for unit
selections/de-selections.
- `none`: no selection is allowed in the graph produced by `girafe`.

All these options can be configured with the following functions:

- `opts_selection`: relative to panel selections
- `opts_selection_key`: relative to legend selections
- `opts_selection_theme`: relative to theme elements selections

The following code is enabling `single` selection in the panel:

```{r}
library(ggiraph)
library(ggplot2)
dataset <- mtcars
dataset$carname <- row.names(dataset)
gg_scatter <- ggplot(dataset, aes(x = disp, y = qsec, tooltip = carname, data_id = carname, color= wt) ) + 
	geom_point_interactive(size=3) +
  theme_minimal()

girafe(ggobj = gg_scatter, 
  options = list(opts_selection(type = "single", only_shiny = FALSE)) )
```

Note that we used `only_shiny = FALSE` so that selections can be seen 
in the vignette but in real shiny applications, you don't have to set this 
argument to FALSE.

### Access the selected values

Three reactive values are available: 

- the selection of the `data_id` of the panel, selected elements will be
captured in the input reactive value `girafeid_selected` (name of the input
id of the reactive output value + `_selected`).
- the selection of the `data_id` of the legend elements, selected elements
will be captured in the input reactive value `girafeid_key_selected` (name
of the input id of the reactive output value + `_key_selected`).
- the selection of the `data_id` of the theme elements, selected elements will
be captured in the input reactive value `girafeid_theme_selected` (name of
the input id of the reactive output value + `_theme_selected`).

```{r eval=FALSE}
input$plot_selected
input$plot_key_selected
input$plot_theme_selected
```

### Update selected values from Shiny 

You can also modify theses values by using the `session$sendCustomMessage`
method with type `girafeid_set` (name of the input id of the reactive output
value + `_set`).

```{r eval=FALSE}
# delete selections
session$sendCustomMessage(type = 'plot_set', message = character(0))
session$sendCustomMessage(type = 'plot_key_set', message = character(0))
session$sendCustomMessage(type = 'plot_theme_set', message = character(0))
```

## Define default selection

Sometimes, it can be useful to pre-select some elements. This can be done by
using the argument `selected` of function `opts_selection` (or
`opts_selection_key` or `opts_selection_theme`).

```{r}
preselection <- dataset$carname[1:5]
girafe(ggobj = gg_scatter, 
  options = list(
    opts_selection(selected = preselection,
                   type = "multiple", only_shiny = FALSE,
                   css = "fill:red;stroke:gray;")))
```



