---
title: "Using ggiraph"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Using ggiraph}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


The **ggiraph** package let R users make **ggplot interactive**. The package is an htmlwidget.
The following graphic is produced by calling `ggiraph()` on a `ggplot` object.

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(ggiraph)
theme_set(theme_minimal())

# geom_point_interactive example
gg_point_1 <- ggplot(mtcars, aes(x = disp, y = qsec, 
		color = wt, tooltip = row.names(mtcars), data_id = row.names(mtcars) ) ) + 
	geom_point_interactive(size=2) + 
  scale_color_gradient(low = "#F3C899", high = "#8C120A") 

# htmlwidget call
ggiraph(code = {print(gg_point_1)}, width_svg = 5, height_svg = 4, width = .75,
        tooltip_extra_css = "padding:2px;background:rgba(70,70,70,0.1);color:black;border-radius:2px 2px 2px 2px;",
        hover_css = "fill:#1279BF;stroke:#1279BF;cursor:pointer;")

```

## Introduction

It extends ggplot2 with new `geom` functions:

```{r echo=FALSE, results='asis'}
geoms_ <- objects(envir = as.environment("package:ggiraph"), pattern = "^geom_([a-zA-Z0-9]*)_interactive$")
htmltools::tags$ul( lapply( geoms_, htmltools::tags$li ) )
```

These understand three aesthetics to let you add interactivity:

* `tooltip`: column of dataset that contains tooltips to be displayed when mouse is over elements. 
* `onclick`: column of dataset that contains javascript function to be executed when elements are clicked. 
* `data_id`: column of dataset that contains id to be associated with elements. **This aesthetic is mandatory 
  when you want to use an hover effect or when you want to enable selection of points in shiny applications.**

Let's prepare a ggplot object with the `mpg` dataset. 

```{r}
library(ggplot2)
library(ggiraph)

head(mpg)
g <- ggplot(mpg, aes( x = displ, y = cty, color = hwy) )
```

### tooltips

The first example shows how to add a tooltip:

```{r message=FALSE}
my_gg <- g + geom_point_interactive(aes(tooltip = model), size = 2) 
ggiraph(code = print(my_gg) )
```

### hover effects

Now let's add an hover effect. **Elements associated with a `data_id` will be animated upon mouse over.**

```{r message=FALSE}
my_gg <- g + geom_point_interactive(
    aes(tooltip = model, data_id = model), size = 2) 
ggiraph(code = print(my_gg), hover_css = "cursor:pointer;fill:red;stroke:red;")
```

The default value of hover css is `hover_css = "fill:orange;"`. [Read more about custom animation effects](#custom-animation-effects)...

> Note that `data-id` can also be reused within a shiny application. 

### Click actions

#### Using onclick

Click actions must be a string column in the dataset containing valid javascript instructions. 

```{r message=FALSE, warning=FALSE}
crimes <- data.frame(state = tolower(rownames(USArrests)), USArrests)
head(crimes)

# create an 'onclick' column
crimes$onclick <- sprintf("window.open(\"%s%s\")",
  "http://en.wikipedia.org/wiki/", as.character(crimes$state) )

gg_crime <- ggplot(crimes, aes(x = Murder, y = Assault, color = UrbanPop )) + 
  geom_point_interactive(
    aes( data_id = state, tooltip = state, onclick = onclick ), size = 3 ) + 
  scale_colour_gradient(low = "#999999", high = "#FF3333")

ggiraph(code = print(gg_crime), hover_css = "fill-opacity:.3;cursor:pointer;")
```

#### Within shiny

When working with shiny, you can use the `data_id` aesthetic to associate points, polygons and other graphical elements with a value that will be available in a reactive context. This makes it possible to click on an element and trigger an action. More details are available in vignette "ggiraph in shiny". Note that in this case, `onclick` should not be used, both onclick and data_id will need the "click" event. [Read more about shiny integration](#shiny-usage)...




## Custom animation effects


With `ggiraph`, you can customize tooltip style and mouse hover effects. This requires usage of `css`.

### Tooltip position

The arguments `tooltip_offx` and `tooltip_offy` are used to offset tooltip position. 

By default the offset is 10 pixels horizontally to the mouse position (`tooltip_offx=10`) and 0 pixels vertically (`tooltip_offx=10`).

```{r message=FALSE}
library(ggplot2)
library(ggiraph)
theme_set(theme_minimal())

dataset <- mtcars
dataset$carname <- row.names(dataset)
gg_point_1 <- ggplot(dataset, aes(x = disp, y = qsec, tooltip = carname, data_id = carname, color= wt) ) + 
	geom_point_interactive(size=3)

# htmlwidget call
ggiraph(code = {print(gg_point_1)}, tooltip_offx = 20, tooltip_offy = -10 )
```


### Tooltip style

The `ggiraph` function has an argument named `tooltip_extra_css`. It can be used to add css declarations to customize tooltip rendering.

> Each css declaration includes a property name and an associated value. Property names and values are separated by colons and name-value pairs always end with a semicolon. For example `color:gray;text-align:center;`. Common properties are : 

> * background-color: background color 
* color: elements color
* border-style, border-width, border-color: border properties
* width, height: size of tooltip
* padding: space around content

Tooltip opacity can be defined with the argument `tooltip_opacity` (default to 0.9).

Let's custom tooltip as:

* italic font 
* no background color

```{r}
tooltip_css <- "background-color:transparent;font-style:italic;"
```

Now print the ggiraph:

```{r message=FALSE}
ggiraph(code = {print(gg_point_1)}, tooltip_extra_css = tooltip_css )
```

Now, let's add a gray rectangle with round borders and a few other details to make it less crude:

```{r}
tooltip_css <- "background-color:gray;color:white;font-style:italic;padding:10px;border-radius:10px 20px 10px 20px;"

ggiraph(code = {print(gg_point_1)}, tooltip_extra_css = tooltip_css, tooltip_opacity = .75 )
```

> Do not surround `tooltip_extra_css` value by curly braces, ggiraph takes care of that.

### Hover effects

Hover effects occur when the mouse is **over elements that have a `data-id` attribute** (resulting from using argument `data_id` in interactive geom functions). It will only modify SVG element rendering when the mouse is over an element. 

Mouse over effects can be configured with the `hover_css` argument in the same way `tooltip_extra_css` is used for customizing tooltip rendering.

css here is relative to SVG elements. SVG attributes are listed [here](https://developer.mozilla.org/fr/docs/Web/SVG/Attribute). Common properties are: 

* fill: background color 
* stroke: color
* stroke-width: border width
* r: circle radius (**no effect if Firefox is used**).

To fill elements in red:

```{r}
ggiraph(code = {print(gg_point_1)}, hover_css = "fill:red;r:10pt;" )
```

### Zoom

You can activate zoom; set `zoom_max` (maximum zoom factor) to a value 
greater than 1. If the argument is greater than 1, a toolbar will 
appear when mouse will be over the graphic. 

Click on the icons in the toolbar to activate or desactivate the zoom.

```{r}
ggiraph(code = print(gg_point_1), zoom_max = 5)
```



## Shiny usage


`ggiraph` is an htmlwidget. It can be used within a Shiny application. In 
shiny, elements associated with `data_id` can be selected and the selection 
(the `data_id` value) is available in the client and the server side of the 
application. 

### The client `ui.R`

Instead of a `plotOutput`, use function `ggiraphOutput`. 

```{r eval=FALSE}
ggiraphOutput("plot")
```

### The server `server.R`

Use function `ggiraphOutput`.

```{r eval=FALSE}
output$plot <- renderggiraph({
    ggiraph(code = print(gg_blahblah) )
  })
```


### Selections

It is possible to work with selected points on a *ggiraph* plot within a Shiny application.

Selection can be of two types: `single` or `multiple`. The ggiraph's parameter `selection_type` will let you specify that. 

```{r eval=FALSE}
output$myplot <- renderggiraph({
    ggiraph(code = print(gg_blahblah), selection_type = "multiple" )
  })
```

The selected points will be captured in the input reactive value `myplot_selected` (name of the input id of the reactive output value + `_selected`):

```{r eval=FALSE}
input$myplot_selected
```


You can also modify theses values by using the `session$sendCustomMessage` method with type `myplot_set` (name of the input id of the reactive output value + `_set`).

```{r eval=FALSE}
# delete selection
session$sendCustomMessage(type = 'myplot_set', message = character(0))
```


### Examples

The package contains Shiny examples available in the `shiny` directory of the package (`system.file("shiny", package = "ggiraph")`).

#### selections usage (server side)

```{r eval=FALSE}
shiny::runApp(appDir = system.file("shiny/crimes", package = "ggiraph"), display.mode = "showcase")
shiny::runApp(appDir = system.file("shiny/cars", package = "ggiraph"), display.mode = "showcase")

# *group* selection
shiny::runApp(appDir = system.file("shiny/iris", package = "ggiraph"), display.mode = "showcase")
```

#### onclick actions (client side).

```{r eval=FALSE}
shiny::runApp(appDir = system.file("shiny/DT", package = "ggiraph"), display.mode = "showcase")
```


