---
title: "Using ggiraph"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Using ggiraph}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
library(knitr)
opts_chunk$set(message=FALSE, fig.width=8, fig.height = 6)
library(usefonts)
library(ggiraph)
library(ggplot2)
add_font("Open Sans")
set_girafe_defaults(fonts = list(sans = "Open Sans"))
theme_set(
  theme_minimal(base_family = "Open Sans")
)
```


```{r echo=FALSE}
addFontHtmlDependency(font = "Open Sans")
```
The **ggiraph** package let R users make **ggplot interactive**. The package 
is an htmlwidget. The following graphic is produced by calling `girafe()` function 
with a `ggplot` object.


```{r}
library(ggiraph)
library(ggplot2)
set_girafe_defaults(
  opts_tooltip = opts_tooltip(
      opacity = 1,
      css = "background-color:#333333;color:white;padding:3px;border-radius:2px;"
    ), 
  opts_hover = opts_hover(css = "fill:magenta;stroke:black;cursor:pointer;r:5px;")
)
gg_point_1 <- ggplot(
  data = mtcars,
  mapping = aes(
    x = disp, y = qsec,
    color = wt, tooltip = row.names(mtcars), data_id = row.names(mtcars)
  )
) +
  geom_point_interactive(size = 3, hover_nearest = TRUE) +
  scale_color_gradient(low = "gray", high = "#8C120A")

x <- girafe(ggobj = gg_point_1)
x
```

## Introduction

It extends ggplot2 with:

- interactive `geom` functions: `geom_point_interactive()`, `geom_col_interactive`, etc.
- interactive `scale` functions: `scale_color_continuous_interactive()`, `scale_fill_manual_interactive()`, etc.
- interactive `facet` functions: `facet_wrap_interactive()` and `facet_grid_interactive()` that both work with `labeller_interactive()`.
- interactive `guide` functions: `guide_colorbar_interactive()`, `guide_colourbar_interactive()`, `guide_legend_interactive()`.
- interactive `theme` elements: `element_line_interactive()`, `element_rect_interactive()`, `element_text_interactive()`, `label_interactive`.

These understand three main aesthetics to let you add interactivity:

* `tooltip`: column of dataset that contains tooltips to be displayed when mouse is over elements. 
* `data_id`: column of dataset that contains id to be associated with elements. **This aesthetic 
is mandatory when you want to use an hover effect or when you want to enable selection of points 
in shiny applications.**
* `onclick`: column of dataset that contains javascript function to be executed when 
elements are clicked. 

Let's prepare a ggplot object with the `mpg` dataset. 

```{r}
library(ggplot2)
library(ggiraph)
g <- ggplot(mpg, aes( x = displ, y = cty, color = hwy) )
```

## tooltips

The first example shows how to add a tooltip:

```{r message=FALSE}
my_gg <- g + geom_point_interactive(aes(tooltip = model), size = 3) 
girafe(code = print(my_gg) )
```

## hover effects

Now let's add an hover effect. **Elements associated with a `data_id` will be 
animated upon mouse over.**

Option `hover_nearest=TRUE` make the hover effect applied on the 
nearest element while moving the mouse. In this case it is 
mandatory to also set the `data_id` parameter.

Note that `data_id` can also be reused within a shiny application as a reactive value. 

```{r message=FALSE}
my_gg <- g + geom_point_interactive(hover_nearest = TRUE,
    aes(tooltip = model, data_id = model), size = 3) 
x <- girafe(code = print(my_gg))
x
```


## Interactive faceting

By using `facet_wrap_interactive()` and `labeller_interactive`,
it is possible to make the strips interactive.

```{r}
dat <- mtcars
dat$carname <- row.names(dat)
p1 <- ggplot(dat, aes(x = mpg, y = qsec, tooltip = carname, data_id = carname)) +
  geom_point_interactive(hover_nearest = TRUE, size = 3) + 
  
  facet_wrap_interactive(ncol = 2,interactive_on = "rect",
    vars(vs),
    labeller = labeller_interactive(aes(
        tooltip = paste("this is vs", vs), data_id = vs
      ))
  )

girafe(ggobj = p1)
```


## Click actions

### Within shiny

When working with shiny, you can use the `data_id` aesthetic to associate
points, polygons and other graphical elements with a value that will be
available in a reactive context. This makes it possible to click on an element
and trigger an action. (Note that in this case, `onclick` should not be used,
both `onclick` and `data_id` will need the "click" event.) 

`data_id` aesthetic is a great and easy way to trigger actions when an element 
such as a point or a region on a map is clicked. The selected `data_id` for elements in panels,
scales and themes are stored in 3 different reactive values. [Read more about shiny
integration](./shiny.html).

### Using onclick

Click actions must be a string column in the dataset containing valid javascript instructions. 

```{r message=FALSE, warning=FALSE}
crimes <- data.frame(state = tolower(rownames(USArrests)), USArrests)
head(crimes)

# create an 'onclick' column
crimes$onclick <- sprintf("window.open(\"%s%s\")",
  "http://en.wikipedia.org/wiki/", as.character(crimes$state) )

gg_crime <- ggplot(crimes, aes(x = Murder, y = Assault, color = UrbanPop )) + 
  geom_point_interactive(hover_nearest = TRUE,
    aes( data_id = state, tooltip = state, onclick = onclick ), size = 3 ) + 
  scale_colour_gradient(low = "#999999", high = "#FF3333")

girafe(ggobj = gg_crime, 
       options = list(
         opts_hover = opts_hover(css = "fill:gray;stroke:black;cursor:pointer;r:6px;")
       ))
```


```{r include=FALSE}
init_girafe_defaults()
```

