---
title: "Using ggiraph"
output: 
  rmarkdown::html_vignette:
    toc: true 
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Using ggiraph}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


The **ggiraph** package let R users make **ggplot interactive**. The package 
is an htmlwidget. The following graphic is produced by calling `girafe()` function 
with a `ggplot` object.

> Note that `girafe()` should replace `ggiraph()` function that has been refactored. 

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(ggiraph)
theme_set(theme_minimal())

# geom_point_interactive example
gg_point_1 <- ggplot(mtcars, aes(x = disp, y = qsec, 
		color = wt, tooltip = row.names(mtcars), data_id = row.names(mtcars) ) ) + 
	geom_point_interactive(size=2) + 
  scale_color_gradient(low = "#F3C899", high = "#8C120A") 

# htmlwidget call
x <- girafe(ggobj = gg_point_1, width_svg = 6, height_svg = 5)
x <- girafe_options(x = x,
    opts_sizing(rescale = FALSE), 
    opts_tooltip(opacity = .8, css = "background-color:gray;color:white;padding:2px;border-radius:2px;"),
    opts_hover(css = "fill:#1279BF;stroke:#1279BF;cursor:pointer;") )
x
```

## Introduction

It extends ggplot2 with:

* interactive `geom` functions: `geom_point_interactive()`, `geom_col_interactive`, etc.
* interactive `scale` functions: `scale_color_continuous_interactive()`, `scale_fill_manual_interactive()`, etc.
* interactive `guide` functions: `guide_colorbar_interactive()`, `guide_colourbar_interactive()`, `guide_legend_interactive()`.

These understand three aesthetics to let you add interactivity:

* `tooltip`: column of dataset that contains tooltips to be displayed when mouse is over elements. 
* `data_id`: column of dataset that contains id to be associated with elements. **This aesthetic 
is mandatory when you want to use an hover effect or when you want to enable selection of points 
in shiny applications.**
* `onclick`: column of dataset that contains javascript function to be executed when 
elements are clicked. 

Let's prepare a ggplot object with the `mpg` dataset. 

```{r}
library(ggplot2)
library(ggiraph)
g <- ggplot(mpg, aes( x = displ, y = cty, color = hwy) )
```

## tooltips

The first example shows how to add a tooltip:

```{r message=FALSE}
my_gg <- g + geom_point_interactive(aes(tooltip = model), size = 2) 
girafe(code = print(my_gg) )
```

## hover effects

Now let's add an hover effect. **Elements associated with a `data_id` will be 
animated upon mouse over.**

```{r message=FALSE}
my_gg <- g + geom_point_interactive(
    aes(tooltip = model, data_id = model), size = 2) 
x <- girafe(code = print(my_gg))
x
```

> Note that `data-id` can also be reused within a shiny application as a reactive value. 

## Click actions

### Within shiny

When working with shiny, you can use the `data_id` aesthetic to associate points, polygons 
and other graphical elements with a value that will be available in a reactive context. 
This makes it possible to click on an element and trigger an action. 
Note that in this case, `onclick` should not be used, both onclick 
and data_id will need the "click" event. [Read more about shiny integration](#shiny-usage)...

### Using onclick

Click actions must be a string column in the dataset containing valid javascript instructions. 

```{r message=FALSE, warning=FALSE}
crimes <- data.frame(state = tolower(rownames(USArrests)), USArrests)
head(crimes)

# create an 'onclick' column
crimes$onclick <- sprintf("window.open(\"%s%s\")",
  "http://en.wikipedia.org/wiki/", as.character(crimes$state) )

gg_crime <- ggplot(crimes, aes(x = Murder, y = Assault, color = UrbanPop )) + 
  geom_point_interactive(
    aes( data_id = state, tooltip = state, onclick = onclick ), size = 3 ) + 
  scale_colour_gradient(low = "#999999", high = "#FF3333")

girafe(ggobj = gg_crime)
```


